---
title: "Sentiments on Safe Injection"
subtitle: "Analysis of Multi-Level Government Content"
author: 
  - Finn Korol
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This paper evaluates the bias of government websites regarding safe injection sites using sentiment analysis. We analyzed website content and identified patterns of sentiment, categorizing them as positive, negative, or neutral. Our findings show varying degrees of bias, with some sites presenting a positive sentiment and others a negative sentiment, using emotionally charged language and negative framing. Biased information can influence public perception and policy decisions, highlighting the importance of accurate and unbiased information. We recommend that government websites provide unbiased information to promote informed decision-making and reduce harm caused by opioid addiction."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```


# Introduction

In recent years, there has been a growing awareness of the public health crisis associated with the opioid epidemic in North America. One approach to mitigate the harm caused by opioid addiction is the establishment of safe injection sites, which provide a medically supervised environment for individuals to use drugs. These sites have been shown to reduce the risk of overdose, decrease the transmission of infectious diseases, and connect individuals with healthcare and social services. Despite their proven effectiveness, safe injection sites remain a controversial and politicized topic in many jurisdictions.

Government websites are a primary source of information for citizens seeking information about safe injection sites. However, it is unclear whether these websites provide unbiased and accurate information. This paper aims to evaluate the bias of government websites regarding safe injection sites using natural language processing techniques, primarily through sentiment analysis.

The use of natural language processing in analyzing text data has become increasingly popular in recent years due to its ability to quickly and accurately process large amounts of text data. Sentiment analysis, in particular, is a widely used method for identifying the underlying emotional tone of text, which can be used to evaluate the positivity or negativity of a particular message or viewpoint. In the context of government websites, sentiment analysis can help identify whether the content is presented in a positive, negative, or neutral manner.

By analyzing the sentiment of government website content related to safe injection sites, this paper aims to shed light on the potential biases that may exist in the information presented to the public. Such biases can have a significant impact on public perception and policy decisions regarding safe injection sites. Therefore, understanding the extent of bias in government websites is essential for informed decision-making and public health policy.



# Data {#sec-data}

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it. 



# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


